---
title: "CCBR1238"
author: "Samantha Sevilla"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
## Define parameters

Set dirs
```{r set_dir, echo=FALSE, warning=FALSE, include=FALSE}
pipeliner_dir="/data/CCBR/projects/ccbr1238/rnaseq/123022_2/"
output_dir="/data/CCBR/projects/ccbr1238/analysis_240303/"
img_dir=paste0(output_dir,"img/")
ipa_dir="/data/CCBR/projects/ccbr1238/IPA_230304/"
GO_dir="/data/CCBR/projects/ccbr1238/analysis_240303/"

dir.create(file.path(output_dir), showWarnings = FALSE)
dir.create(file.path(img_dir), showWarnings = FALSE)
```

Set parameters
```{r}
species_in="Mus musculus"
log2_cutoff=1.3
fdr_cutoff=0.05
padj_cutoff=0.05
csid="ccbr1238"
sample_consensus_threshold=0.1
read_minimum_threshold=10

# set list of databases to use for GO and KEGG analysis
# http://www.gsea-msigdb.org/gsea/msigdb/collections.jsp
db_list=c("C5:MF","C5:BP","C5:CC","C2:KEGG")

# Minimal size of a gene set to test. All pathways below the threshold are excluded.
minSize_gene_set=15
```

Set contrasts
```{r set_contrasts, echo=FALSE, warning=FALSE}
fpath=paste0(pipeliner_dir,"groups.tab")
groups_df=read.csv(fpath,sep="\t",header=FALSE)
colnames(groups_df)=c("replicate","group","replicate2")
head(groups_df)  

contrasts_list=c("SH1_0pt5mM-SCR_0pt5mM",
                 "SH4_0pt5mM-SCR_0pt5mM",
                 "SH2_0pt5mM-SCR_0pt5mM",
                 "SCR_4mM-SCR_0pt5mM",
                 "SCR_0pt5mM-SCR_4mM",
                 "SH1_0pt5mM-SCR_4mM",
                 "SH4_0pt5mM-SCR_4mM")
```

Set pkgs
```{r set_pkg, echo=FALSE, warning=FALSE, include=FALSE}
if (species_in == "Homo sapiens"){
  species_db="org.Hs.eg.db"
  species_short="hsa"
} else if (species_in == "Mus musculus"){
  species_db="org.Mm.eg.db"
  species_short="mmu"
} else{
    print("Species not approved")
}

# package list
list.of.packages=c("BiocManager","tidyr","DT","RUVSeq","RColorBrewer","textshape",
                   "pheatmap","grid","dplyr","EnhancedVolcano","edgeR","DESeq2","ggplot2","plotly",
                   "msigdbr","clusterProfiler","ggpubr","ggridges","factoextra",
                   "gplots","ReactomePA",species_db,"pander","ggVennDiagram","fgsea","stringr",
                   "gridExtra","tidyverse","scales","cowplot")

#install as needed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)

# source functions
source("/data/CCBR/projects/ccbr1238/scripts/ccbr1238_source_code.R")

# load packages
invisible(lapply(list.of.packages, library, character.only = TRUE))

#negate
`%ni%` <- Negate(`%in%`)
```

## Run Analysis

### HEATMAP

Heat map of the following genes in these 4 samples. Please take 2 samples from each condition. So there will be a total of 8 columns of heat maps. Keep cut off >1.3

These are the genes: ATF4, SLC7A1, SLC7A3, SLC7A5, SLC7A11, SLC3A2, SLC6A9, SLC38A1, SLC38A2, SLC38A7, SLC38A9, SLC1A4, SLC1A5

Expectation: Here, at SCR 0.5 mM_1 and SCR 0.5 mM_2, the expression of these genes should be upregulated in comparison to SCR 4 mM_1, SCR 4mM_2. In SH1 0.5 mM_1, SH2 0.5 mM_1, SH4 0.5 mM_1, and SH5 0.5 mM_1, the expression of these genes should be down.

Notes: Because these are normalized, scaled, values, there is no FC to use as a cutoff. FC only can be applied when a comparison is being made.

```{r heatmap, echo=FALSE}
# sample selection
sample_list=c("SCR_4mM_1","SCR_4mM_2","SCR_0pt5mM_1","SCR_0pt5mM_2","SH1_0pt5mM_1",
              "SH1_0pt5mM_2","SH2_0pt5mM_1","SH2_0pt5mM_2")
print(subset(groups_df,replicate %in% sample_list)[,c("replicate","group")])

# gene list
gene_list=c("Slc1a4","Slc1a5","Slc38a1","Slc38a2","Slc38a7","Slc38a9","Slc3a2",
            "Slc6a9","Slc7a1","Slc7a2","Slc7a3","Slc7a5","Slc7a7","Slc7a11")
#sample_list=sample_list;gene_list=gene_list;scale_flag="ON";name_flag="ON"
MAIN_REPLICATE_HEATMAPS(sample_list=sample_list,
                        gene_list=gene_list,
                        scale_flag="ON",
                        name_flag="ON")

# gene list
gene_list=c("Atf4","Phgdh","Psph","Psat1","Aldh1l2","Mthfd2","Shmt2","Mthfd1l","Cth","Bcat1",
            "Gpt2","Gstp1","Gstp2","Gstk1","Gpx2","Chac1")
MAIN_REPLICATE_HEATMAPS(sample_list=sample_list,
                        gene_list=gene_list,
                        scale_flag="ON",
                        name_flag="ON",
                        pre="ON")

gene_list=c("Atf4","Phgdh","Psph","Psat1","Aldh1l2","Mthfd2","Shmt2","Mthfd1l","Cth","Bcat1",
            "Gpt2","Gstp1","Gstp2","Gpx2","Chac1")
MAIN_REPLICATE_HEATMAPS(sample_list=sample_list,
                        gene_list=gene_list,
                        scale_flag="ON",
                        name_flag="ON")

gene_list=c("Atf4","Phgdh","Psph","Psat1","Aldh1l2","Mthfd2","Shmt2","Mthfd1l","Cth","Bcat1",
            "Gpt2","Slc6a9","Slc7a11","Gstp1","Gstp2","Gpx2","Chac1")
MAIN_REPLICATE_HEATMAPS(sample_list=sample_list,
                        gene_list=gene_list,
                        scale_flag="ON",
                        name_flag="ON")

```

Plot pathways by pvalue
```{r}
ipa_pathways=read.csv("/data/CCBR/projects/ccbr1324/analysis/ipa_240303/pathway_analysis.csv")
colnames(ipa_pathways)=c("pathway","pvalue","ratio","zscore","molecules")
head(ipa_pathways)

pathway_list=c("Tumor Microenvironment Pathway","Granzyme A Signaling","IL-10 Signaling",
               "PIP3 activates AKT signaling","IL-15 Signaling","mTOR Signaling",
               "Class I MHC mediated antigen processing and presentation",
               "Neutrophil Extracellular Trap Signaling Pathway",
               "IL-8 Signaling")
library(forcats)
library(scales)
ipa_pathways_sub=subset(ipa_pathways,pathway %in% pathway_list)

scale_factor <-2
ipa_pathways_sub$ratioPerc=(ipa_pathways_sub$ratio)*100

ipa_pathways_sub %>%
    ggplot(aes(x = pathway))+
    geom_bar(aes(y = ratioPerc,fill=molecules),colour="red", stat="identity")+
    geom_point(aes(y = pvalue,size=pvalue,color=""),shape=20, cex=8)+ # create dummy points so pvalue legend symbol appears
    geom_point(aes(y = pvalue,size=pvalue),shape=20, cex=8,color="blue")+
    scale_color_manual(name = "log10(pvalue) x \nActivation Z Score",values = c(" " = "blue")) + # add legened for pvalue point
    scale_fill_brewer(direction = -1, palette = "Pastel2")+ # change palette for bars
    coord_flip()+ # flip x and y axis
    scale_y_continuous(sec.axis = sec_axis(~ .*1, #create a secondary "x" axis for the log10(pvalue data)
                                           labels = number_format(scale=scale_factor), 
                                           name="log10(pvalue) x sign of Activation Z Score"))+
    xlab("Select Pathway")+
    ylab("Gene Ratio %")+
    theme_bw() + # remove grey background
    theme( axis.line.x.top = element_line(color = "blue"),axis.ticks.x.top= element_line(color = "blue"), # change top x axis color to match pvalues
           axis.text.x.top = element_text(color = "blue"),axis.title.x.top=element_text(colour="blue"))+
    theme( axis.line.x.bottom = element_line(color = "red"),axis.ticks.x.bottom= element_line(color = "red"), # change bottom x axis color to match z scores
           axis.text.x.bottom = element_text(color = "red"),axis.title.x.bottom=element_text(colour="red"))+
    ggtitle(paste0("Select Pathways")) + # add title
    theme(plot.title = element_text(size=22, face="bold")) + #  increase title
    theme(legend.key=element_rect(fill=NA)) +# remove grey from pvalue background
    labs(color="log10(pvalue) x \nActivation Z Score",fill="Activation Z Score")+ # change the order of the legend so log10 is on top
    guides(fill  = guide_legend(order = 2),color = guide_legend(order = 1))+
    theme(legend.position = "none") # remove legend
```

