---
title: "CCBR1238"
author: "Samantha Sevilla"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
## Background information

### Set dirs
```{r set_dir, echo=FALSE, warning=FALSE, include=FALSE}
pipeliner_dir="~/../../Volumes/ccbr1238/rnaseq/123022_2/"
output_dir="~/../../Volumes/ccbr1238/analysis_230221/"

dir.create(file.path(output_dir), showWarnings = FALSE)
```

### Set thresholds
```{r}
species_in="Homo sapiens"
log2_cutoff=1.5
fdr_cutoff=0.05
```

### Set pkgs
```{r set_pkg, echo=FALSE, warning=FALSE, include=FALSE}
if (species_in == "Homo sapiens"){
  species_db="org.Hs.eg.db"
} else if (species_in == "Mus musculus"){
  species_db="org.Mm.eg.db"
} else{
    print("Species not approved")
}
# package list
list.of.packages=c("BiocManager","tidyr","DT","RUVSeq","RColorBrewer","textshape",
                   "pheatmap","grid","dplyr","EnhancedVolcano","edgeR","DESeq2","ggplot2","plotly",
                   "msigdbr","clusterProfiler","ggpubr","ggridges","factoextra",
                   "gplots","ReactomePA",species_db,"pander","ggVennDiagram")

#install as needed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)

# source functions
source("~/../../Volumes/ccbr1238/scripts/ccbr1238_source_code.R")

# load packages
invisible(lapply(list.of.packages, library, character.only = TRUE))

#negate
`%ni%` <- Negate(`%in%`)

```

### Set contrasts
```{r set_contrasts, echo=FALSE, warning=FALSE, include=FALSE}
fpath=paste0(pipeliner_dir,"groups.tab")
groups_df=read.csv(fpath,sep="\t",header=FALSE)
colnames(groups_df)=c("replicate","group","replicate2")
head(groups_df)  

contrasts_list=c("SCR_4mM-SCR_0pt5mM",
                 "SH1_0pt5mM-SCR_0pt5mM",
                 "SH2_0pt5mM-SCR_0pt5mM",
                 "SH3_0pt5mM-SCR_0pt5mM",
                 "SH4_0pt5mM-SCR_0pt5mM",
                 "SH5_0pt5mM-SCR_0pt5mM",
                 "SH6_0pt5mM-SCR_0pt5mM")
```

## Quality Assessment
### BEFORE NORMALIZATION

Filtered data findings:
- SCR samples under different conditions cluster separately from one another.
- SH samples by group cluster separately: GCN1 KD (SH1-3) and ATF4 KD (SH4-6). 
    - The two groups are not as uniform within SH types: SH1-3 cluster tightly, SH4-5 cluster tightly, SH6 clusters alone.
- All replicates cluster tightly together.

```{r qc_f, echo=FALSE, warning=FALSE}
group_list=c("SCR_4mM","SCR_0pt5mM",
             "SH1_0pt5mM","SH2_0pt5mM",
             "SH3_0pt5mM","SH4_0pt5mM",
             "SH5_0pt5mM","SH6_0pt5mM")
MAIN_PREP_QC_CORE(groups_in=group_list,
                  stage_in="before_norm")
```

### AFTER NORMALIZATION:
Upper quantile normalization findings:
- This did not provide much correction over the filtered counts.
```{r qc_upper, echo=FALSE,warning=FALSE}
group_list=c("SCR_4mM","SCR_0pt5mM",
             "SH1_0pt5mM","SH2_0pt5mM",
             "SH3_0pt5mM","SH4_0pt5mM",
             "SH5_0pt5mM","SH6_0pt5mM")
MAIN_PREP_QC_CORE(groups_in=group_list,
                  stage_in="upper_quant")
```

DESEQ2 normalization findings:
- This provided slightly better correction than upper quantile, but also did not significantly deviate from the filtered counts. This method is suggested for further analysis.

```{r qc_deseq, echo=FALSE,warning=FALSE, message=FALSE}
group_list=c("SCR_4mM","SCR_0pt5mM",
             "SH1_0pt5mM","SH2_0pt5mM",
             "SH3_0pt5mM","SH4_0pt5mM",
             "SH5_0pt5mM","SH6_0pt5mM")
MAIN_PREP_QC_CORE(groups_in=group_list,
                  stage_in="DESEQ")
```

### Run DEG Analysis
Find differential differences between each of the contrasts
```{r deg_prim, echo=FALSE,warning=FALSE, message=FALSE}
for (contrast_id in contrasts_list){
  PRIMARY_DIFFERENTIAL(contrast_id)
}
```

Find the overlap between contrasts: SH1, SH2, SH3
```{r deg_o1, echo=FALSE,warning=FALSE, message=FALSE}
compare_list=c(contrasts_list[2],
               contrasts_list[3],
               contrasts_list[4])
subtitle="SH1,SH2,SH3"
COMPARATIVE_VENN_DIAGRAM(compare_list_in = compare_list,
                         subtitle_in = subtitle)
```

Find the overlap between contrasts: SH4, SH5, SH6
```{r deg_o2, echo=FALSE,warning=FALSE, message=FALSE}
compare_list=c(contrasts_list[5],
               contrasts_list[6],
               contrasts_list[7])
subtitle="SH5,SH6,SH7"
COMPARATIVE_VENN_DIAGRAM(compare_list_in = compare_list,
                         subtitle_in = subtitle)
```