---
title: "ccbr1311_240130"
author: "Samantha Sevilla"
date: "2024-01-30"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE, include=FALSE, message=FALSE, warn=FALSE}
parent_dir="/data/CCBR/projects/ccbr1324/"
rna_dir=paste0(parent_dir,"analysis/renee_240214/")
output_dir=paste0(parent_dir,"analysis/ranalysis_240130/")
img_dir=paste0(output_dir,"img/")
ipa_dir=paste0(parent_dir,"analysis/ipa_240220")
dir.create(file.path(output_dir), showWarnings = FALSE)
dir.create(file.path(img_dir), showWarnings = FALSE)

padj_cutoff=0.05
log2fc_set=1.1
species_in="Mus musculus"
org_db="org.Mm.eg.db"
species_short="mm"

```

```{r libs, echo=FALSE, include=FALSE, message=FALSE, warn=FALSE}
# source functions
source("functions.R")

# install as needed, load packages
list.of.packages=c("RColorBrewer","ggfortify","DT","ggVennDiagram",
                   "EnhancedVolcano","org.Mm.eg.db","dplyr","ggrepel","msigdbr",
                   "ggpubr","EDASeq","PCAtools","dplyr","pheatmap","DESeq2","clusterProfiler",
                   "biomaRt","tidyverse","scales","cowplot")
handle_packages(list.of.packages)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene::TxDb.Mmusculus.UCSC.mm10.knownGene

#negate
`%ni%` <- Negate(`%in%`)
```

```{r metadata, include=FALSE, message=FALSE, warn=FALSE}
rna_meta_file=paste0(parent_dir,"docs/rna_meta.csv")
rna_meta=read.csv(rna_meta_file)
rownames(rna_meta)=rna_meta$filename
head(rna_meta)

exclude_list=c("")
print(paste0("The following samples are excluded: ", exclude_list))
rna_meta=subset(rna_meta, sampleid %ni% exclude_list)
DT::datatable(rna_meta)

#set colors
colors <- brewer.pal(6, "Set2")
g=unique(rna_meta$group)
g=as.factor(g)#set pheno data

m=unique(rna_meta$method)
m=as.factor(m)#set pheno data
```

# RNA Analysis {.tabset}

```{r norm_genes, include=FALSE, message=FALSE, warn=FALSE}
# bring in counts
counts_file=paste0(rna_dir,"DEG_ALL/RSEM.genes.expected_count.all_samples.txt")
counts_df=read.csv(counts_file,sep="\t")
colnames(counts_df)=gsub("X","",colnames(counts_df))
colnames(counts_df)=gsub("_001","",colnames(counts_df))
head(counts_df)

#######################################
## filter and convert to matrix
filtered_df=filter_low_counts(counts_df)
filtered_mat=counts_dat_to_matrix(filtered_df)
```


## Quality {.tabset}

The first two plots show the data before and after normalization. PCA plots are difficult to review given the sample size.

PCA's were regenerated larger, using a different tool to correlate variance. Here you can see the overall clustering of samples by group and within groups.

The Eigencor plot shows effect of group, harvestID and biological replicates on principal components. Significant effects were observed group (PC1) and harvestID and biologicalReplicates (PC2). The scale bar represents Eigenvalues (− 1 to 1, with 0 having no correlation). For this analysis the lower 10% of variables based on variance were removed.

### Method {.tabset}

Due to processing differences, method was first reviewed. 

```{r method, include=FALSE, message=FALSE, warn=FALSE}
#######################################
# DESEQ2
dds <- DESeq2::DESeqDataSetFromMatrix(countData = filtered_mat,
                                colData = rna_meta[,c("sampleid","method")],
                                design = ~ method)
dds <- DESeq2::DESeq(dds)
res <- DESeq2::results(dds)

## pull data
Normalized_counts_matrix=as.data.frame(counts(dds, norm=TRUE))
Normalized_counts_matrix$gene_id=rownames(Normalized_counts_matrix)
Normalized_counts_matrix=Normalized_counts_matrix %>%
    tidyr::separate(gene_id,sep="[.]",c("ENSEMBL","ID")) 
fpath=paste0(output_dir,"normalized_counts.csv")

PLOT_RLE_PCA(filtered_mat[,c(2:ncol(filtered_mat))],"Fig. Before Normalization\nAll Samples",g)
PLOT_RLE_PCA(Normalized_counts_matrix[,c(1:(ncol(Normalized_counts_matrix)-2))],"Fig. After Normalization\nAll Samples",m)

PLOT_PCA(dds,rna_meta,"","All Samples","raw","method")
PLOT_PCA(dds,rna_meta,"","All Samples","normalized","method")

# norm read differences by method
vst <- assay(vst(dds))
rownames(rna_meta)=rna_meta$sampleid
rna_meta=dplyr::slice(rna_meta,match(as.vector(colnames(vst)), rna_meta$sampleid))
p <- PCAtools::pca(vst, metadata = rna_meta, removeVar = 0.1)
PCAtools::eigencorplot(p,
    metavars = c('group','method'))

# run with both variables
dds <- DESeq2::DESeqDataSetFromMatrix(countData = filtered_mat,
                                colData = rna_meta[,c("sampleid","method","group")],
                                design = ~ method + group)
dds <- DESeq2::DESeq(dds)
res <- DESeq2::results(dds)
select <- order(rowMeans(counts(dds,normalized=TRUE)),
                decreasing=TRUE)[1:20]
df <- as.data.frame(colData(dds)[,c("method","group")])
pheatmap(assay(vst(dds))[select,], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=FALSE, annotation_col=df)
```

### Group {.tabset}

Evaluation of the methods parameters led to exclusion of all `std` samples. Normalization is performed without these samples, to determine the effects of groups.

```{r}
# subset data to excluse any std samples
exclude_list=c("13_IMPACT_8_std","2_CNTRL_2_std","5_CNTRL_5_std","6_CNTRL_6_std","7_IMPACT_1_std","8_IMPACT_2_std")
print(paste0("The following samples are excluded: ", exclude_list))
rna_meta_sub=subset(rna_meta, sampleid %ni% exclude_list)
DT::datatable(rna_meta_sub)
filtered_mat_sub=filtered_mat[,c(rna_meta_sub$sampleid)]

# DESEQ2
dds <- DESeq2::DESeqDataSetFromMatrix(countData = filtered_mat_sub,
                                colData = rna_meta_sub[,c("sampleid","group")],
                                design = ~ group)
dds <- DESeq2::DESeq(dds)
res <- DESeq2::results(dds)

## pull data
Normalized_counts_matrix=as.data.frame(counts(dds, norm=TRUE))
Normalized_counts_matrix$gene_id=rownames(Normalized_counts_matrix)
Normalized_counts_matrix=Normalized_counts_matrix %>%
    tidyr::separate(gene_id,sep="[.]",c("ENSEMBL","ID")) 

PLOT_RLE_PCA(Normalized_counts_matrix[,c(1:(ncol(Normalized_counts_matrix)-2))],"Fig. After Normalization\nAll Samples",g)
PLOT_PCA(dds,rna_meta_sub,"","All Samples","normalized","group")
```

Samples are still unbalanced. A subset of remaning samples were included.
```{r}
# subset data to excluse any std samples
exclude_list=c("13_IMPACT_8_std","2_CNTRL_2_std","5_CNTRL_5_std",
               "6_CNTRL_6_std","7_IMPACT_1_std","8_IMPACT_2_std","10_IMPACT_4","11_IMPACT_5",
               "1_CNTRL_1","13_IMPACT_8","4_CNTRL_4","3_CNTRL_3")
print(paste0("The following samples are excluded: ", exclude_list))
rna_meta_sub=subset(rna_meta, sampleid %ni% exclude_list)
DT::datatable(rna_meta_sub)
filtered_mat_sub=filtered_mat[,c(rna_meta_sub$sampleid)]

# DESEQ2
dds <- DESeq2::DESeqDataSetFromMatrix(countData = filtered_mat_sub,
                                colData = rna_meta_sub[,c("sampleid","group")],
                                design = ~ group)
dds <- DESeq2::DESeq(dds)
res <- DESeq2::results(dds)

## pull data
Normalized_counts_matrix=as.data.frame(counts(dds, norm=TRUE))
Normalized_counts_matrix$gene_id=rownames(Normalized_counts_matrix)
Normalized_counts_matrix=Normalized_counts_matrix %>%
    tidyr::separate(gene_id,sep="[.]",c("ENSEMBL","ID")) 

PLOT_RLE_PCA(Normalized_counts_matrix[,c(1:(ncol(Normalized_counts_matrix)-2))],"Fig. After Normalization\nAll Samples",g)
PLOT_PCA(dds,rna_meta_sub,"","Subset Samples","normalized","group")
```

```{r anno, echo=FALSE}
# annotation
annotated_df=as.data.frame(res)
annotated_df$ENSEMBL=rownames(annotated_df)
annotated_df=annotated_df %>%
    tidyr::separate(ENSEMBL,sep="[.]",c("ENSEMBL","ID")) 
mouse = useMart("ensembl", dataset = "mmusculus_gene_ensembl")
mouse_res <- getBM(attributes = c("ensembl_gene_id", "external_gene_name","chromosome_name",
                            'strand','transcript_start','transcript_end'), 
             mart = mouse)
annotated_df=merge(annotated_df,mouse_res,by.x="ENSEMBL",by.y="ensembl_gene_id",all.x=TRUE)

# remove dup ensembl
annotated_df=annotated_df[!duplicated(annotated_df$ENSEMBL), ]
head(annotated_df)
colnames(annotated_df)=gsub("external_gene_name","SYMBOL",colnames(annotated_df))
```

```{r sig}
# determine sig
log2_cutoff=1.1
padj_cutoff=0.05
annotated_df$significance="N"
annotated_df$significance[abs(annotated_df$log2FoldChange)>log2_cutoff & annotated_df$padj<padj_cutoff ]="Y"

# print out result
print(paste0("--the number of significant genes found: ",nrow(subset(annotated_df,significance=="Y"))))
  
# write out DEG
fpath=paste0(output_dir,"DEG_sigs.csv")
write.csv(annotated_df,fpath,row.names=FALSE)
```

```{r gsea}
# gather gene list
ref_list=c("H","C1","C5:CC","C5:BP","C5:MF","C2:WIKIPATHWAYS","C2:REACTOME","C2:BIOCARTA")
for (ref in ref_list){
  # create a gene list with the necessary geneID formatting
  gl=deg2geneList(annotated_df,ref)
  
  RUN_CLUSTERPROFILER(ref)
}
```


```{r}
# read in ipa results
fpath=paste0(parent_dir,"/analysis/ipa_240220/upstream_regulators.txt")
ipa_results=read.csv(fpath,sep="\t")
mol_list=c("cytokine","transcription regulator","growth factor","kinase","transmembrane receptor")

nmol_list=c("1","2","3")
for (nmol in nmol_list){
  PLOT_REGULATORS(nmol)
}
PLOT_REGULATORS<-function(nmol){
  print(paste0("--molecules 1 to ",nmol))
  
  # create top mol list
  mol_num=c(1:nmol)
  top_df=create_top_list(ipa_results,mol_num)
    
  # create plot
  create_zscore_pvalue_plot(top_df,mol_num)
    
  #create DT
  select_cols=c("Molecule.Type","Upstream.Regulator","Activation.z.score","p.value.of.overlap","logpvalue")
  out_df=top_df[,select_cols]
  out_df$logpvalue=signif(out_df$logpvalue, 4)
  colnames(out_df)=c("Molecule.Type","Upstream.Regulator","Activation.z.score","pvalue","log10.pvalue")
  DT::datatable(out_df,caption = paste0("Top ", max(mol_num), " Upstream Regulators by Molecule Type"))
}
```


